---
title: "3_CMV_Inflamamtion"
author: "Sergio Andreu-Sanchez"
date: "5/23/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## CMV Omics - Script 3: CMV effects in cytokine production and pro-CVD markers

*R Markdown* of third script. Read data of LLD CMV seropositivity predictions (from script 1), blood cytokines/adipokines and CVD proteomics panel.


```{r libraries, message=FALSE, warning=FALSE, results='hide'}
Location_dir = "~/Documents/GitHub/CMV_Associations/"

library(MASS) #For LDA_function
library(tidyverse)
library(caret) #For RF_function
library(MetBrewer)
library(pROC)
library(glmnet)
library(glmnetUtils) 
library(patchwork)
library(ggrepel)
library(ggforce)
library(readxl)


source(paste0(Location_dir,"scripts/Functions.R"))
source(paste0(Location_dir,"scripts/Functions_multivariate.R"))
```

Open data and process.
1. Covariates including CMV prediction
2. Cytokines/Adipokines
3. Proteomics CVD OLINK panel
4. Cell predictions (CLR/ALR transform and PCA for dimensionality reduction)


```{r Read data,  message=FALSE, warning=FALSE, results='hide'}
#CMV predictions
Covariates = read_tsv(paste0(Location_dir,"Results/LLD_CMV_profile.tsv")) %>% filter(TimePoint == "Baseline")
#Cytokines
Phenotypes <- read_delim(paste0(Location_dir,"Data/Merged_Phenotypes_LLD.csv"), delim="|")
Phenotypes %>% dplyr::select(ID,Citrullin, IL.1.Beta, IL.6, IL.8, IL.10, IL.12P70, `IL-18`, `IL-18bpx`, TNF.Alpha, `Leptin `, Resistin, `Adiponectin `, `AAT `, `hs-CRP`) %>% drop_na() %>% rename ( "IL_18" = `IL-18` , "AAT"=`AAT `, "Adiponectin"=`Adiponectin `, "Leptin"=`Leptin `, "hsCRP"=`hs-CRP`, "IL_18bpx" =`IL-18bpx`) -> Cytokines
#Proteomics uncorrected
read_tsv(paste0(Location_dir,"Data/CVD3_olinkNormal_1447_LLDsamples_t_ProtNames.txt")) -> Proteomics
Proteomics %>% t() %>% as.data.frame() %>%  rownames_to_column("ID") %>% as_tibble() %>% filter(!ID == "Protein") %>% `colnames<-`(c("ID", Proteomics$Protein))  -> Proteomics
Proteomics %>% select(-CCL22) -> Proteomics

#Cell predictions
all_cell_types = read_tsv(paste0(Location_dir,"Data/BIOS_cell_types_DeconCell_2019-03-08.LLD_subset.txt")) -> all_cell_types
#format
all_cell_types %>% gather(key = var_name, value = value, 2:ncol(all_cell_types)) %>% spread_(key = names(all_cell_types)[1],value = 'value') %>% rename(ID = var_name) -> all_cell_types
# negative values to 0, recalculate to 100 and ALR transformation using monocytes as denominator
Clean_and_ALR(all_cell_types) -> cell_matrix_alr
Clean_and_CLR(all_cell_types) -> cell_matrix_clr
cell_matrix_xlr = cell_matrix_clr

# logs of 0 give infinite, make NA and remove samples
cell_matrix_xlr[sapply(cell_matrix_xlr, is.infinite)] <- NA
cell_matrix_xlr %>% drop_na() -> cell_matrix_xlr
#Make a PCA to control for cell composition
res.pca <- prcomp(dplyr::select(cell_matrix_xlr,-ID), scale = TRUE)
Variability = 100* (res.pca$sdev/sum(res.pca$sdev))
PCs = as_tibble(res.pca$x) %>% mutate(ID = cell_matrix_xlr$ID)
PCs %>% dplyr::select(c(ID, PC1, PC2,PC3,PC4,PC5,PC6,PC7,PC8,PC9,PC10)) -> Top_PCs
```



```{r Some helper functions}
#Null: denominator uses the null distribution to estimate discoveries
Compute_FDR_null =  function(P, Permutation_results,fdr_threshold = 0.05 ){
  # calculate FDR for each observed P-value
  fdr_values <- sapply(P, function(p) {
    # count number of false positives (i.e., permuted P < observed P)
    false_positives <- sum(Permutation_results <= p)
    # calculate FDR
    fdr <- false_positives / sum(Permutation_results <= fdr_threshold)
    # ensure FDR is between 0 and 1
    fdr <- ifelse(fdr < 0, 0, ifelse(fdr > 1, 1, fdr))
    return(fdr)
  })
  return(fdr_values)
} 
#FDR2: Denominator uses the number of discoveries in real P. N FD is corrected by the number of permutations
Compute_FDR2 =  function(P, Permutation_results,Permutations = 100 ){
  Order = order(P)
  tibble(Position = Order, FDR = NA) -> values
  fdr_values = c()
  for ( i in seq(P) ){
    Threshold = P[which(Order == i)]
    FDR_v = sum(Permutation_results <= Threshold) / Permutations / i
    if (FDR_v > 1 ){ FDR_v = 1 }
    values %>% mutate(FDR = ifelse(Position == i, FDR_v, FDR  )) -> values
  }
  return(values$FDR)
}  
Basic_association2 = function(DF=For_association, Cyto=Cytokines, Cell_count_correction=T){
  Cyto_association = tibble()
  for (C in colnames(Cyto)){
      if (C ==  "ID"){ next }
      if (Cell_count_correction == F){
        Formula = as.formula(paste0(C, " ~ Age+Sex+CMV_prediction"  ) )
      } else {
        Formula = as.formula(paste0(C, " ~ Age+Sex+CMV_prediction+PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC10"  ) )
      }
      lm(Formula,DF) -> M
      M %>% summary() -> S
      S$coefficients["CMV_prediction",] %>% t() %>% as_tibble() %>% mutate(Cytokine = C, Sample_n = length(M$residuals)) -> S
      Cyto_association = rbind(Cyto_association, S)
}
 return(Cyto_association) 
}  

```

Cytokines analyses.
Check distribution of cytokines. Identify if log10 is needed to reduce distance. Observe if 0 inflation is a problem.
Association with CMV status (linear model). Permutations to estimate FDR.

```{r Explore cytokine distributions and association,  message=FALSE, warning=FALSE}

Cytokines %>% gather(Cytokine, Concentration ,2:dim(Cytokines)[2]) %>% ggplot(aes(x= log10(Concentration) )) + geom_histogram() + theme_bw() + facet_wrap(~Cytokine, scales="free")

set.seed(76)

Cytokines %>% dplyr::select(-ID) %>% apply(2,log10) %>% as_tibble() %>% apply(2, function(x){ x[x==min(x)] = NA ; return(x) } )  %>% as_tibble() %>% mutate(ID = Cytokines$ID) -> Add_cytokines
left_join( dplyr::select(Covariates, c(ID, Age, Sex, CMV_prediction)) , Add_cytokines, by="ID") -> For_association
left_join(For_association, Top_PCs, by="ID") -> For_association #If cell count correction is wanted
For_association[sapply(For_association, is.infinite)] <- NA


Permutations_3 = function( F_A, Cyto, Real_P, N_Permutations=100 ){
  #We will calculate FDR values through permutation
  Null_distribution = c()
  for (Permutation_n in seq(N_Permutations) ){
    #print( paste0("Permutation: ", Permutation_n ) )
    
    F_A_permuted = F_A
    F_A_permuted$CMV_prediction =  sample(F_A_permuted$CMV_prediction, size = dim(F_A_permuted)[1], replace=F )

    Basic_association2(F_A_permuted, Cyto)  -> Permuted_results
    Null_distribution = c(Null_distribution, Permuted_results$`Pr(>|t|)`)
  }
  
  FDRs = Compute_FDR2(Real_P, Null_distribution , N_Permutations)
  
  return(FDRs)
}


Basic_association2(For_association, Cytokines) %>% arrange(`Pr(>|t|)`) -> Cyto_association
Permutations_3(For_association, Add_cytokines, Cyto_association$`Pr(>|t|)`, 1000) -> FDR_p
Cyto_association %>% mutate(FDR_perm = FDR_p) -> Cyto_association
print(Cyto_association)


write_tsv(Cyto_association, paste0(Location_dir, "Results/CMV_Cytokines.tsv"))

```
Residuals of 0-inflated distributions dont look fine. I believe we might be seing some lower limit detention problems (min of TNF-alpha is 0.16 in many samples, 0.21 for IL.1.Beta ...). 


We find that CMV infection is significantly associated with higher levels of IL_18BP (0.0409, P=0.00166, FDR_bh=0.0232, FDR_1000=0.031). Tolerogenic response. This cytokine's distribution looks ok.

Next, we will explore the OLINK CVD proteomic panel.

```{r Explore proteomics distributions,  message=FALSE, warning=FALSE}

###PREPARE DATA #####
#Olink data is already log10 transformed
Proteomics %>% dplyr::select(-ID) %>% apply(2, function(x){ as.numeric(x) } ) %>% as_tibble() %>% mutate(ID = Proteomics$ID) %>% left_join(dplyr::select(Covariates, c(ID, CMV_prediction, Age, Sex)) , .) %>% drop_na()  -> DF_p_uncorrected


#Not-adjusted data
DF_p_uncorrected[sapply(DF_p_uncorrected, is.infinite)] <- NA
DF_p_uncorrected %>% drop_na() -> DF_p_uncorrected
DF_p_uncorrected %>% dplyr::select(-c(CMV_prediction, Sex, ID)) %>% apply(2, scale) %>% as_tibble() %>% mutate(CMV_prediction = DF_p_uncorrected$CMV_prediction, Sex= DF_p_uncorrected$Sex, ID=DF_p_uncorrected$ID) -> DF_p_uncorrected
Saved_IDs_un = DF_p_uncorrected$ID
#Add cell composition covariates
left_join(DF_p_uncorrected, Top_PCs) %>% dplyr::select(-ID) %>% drop_na() -> DF_p_uncorrected_PCs
DF_p_uncorrected %>% dplyr::select(-ID) -> DF_p_uncorrected

DF_p_uncorrected %>% select(-c(Age, Sex, CMV_prediction)) %>% gather(Protein, Concentration) %>% ggplot(aes(x=Concentration)) + geom_histogram() + theme_bw() + facet_wrap(~Protein)

DF_p_uncorrected  %>% mutate(CMV_prediction = as.numeric(CMV_prediction)) %>% cor() -> For_plot
For_plot %>% pheatmap::pheatmap(labels_row = make_bold_names(For_plot, rownames, c("CMV_prediction")), labels_col = make_bold_names(For_plot, colnames, c("CMV_prediction") ), color = met.brewer(name="Cassatt2", type="continuous") )
```


```{r Association proteomics,  message=FALSE, warning=FALSE}

Assoc_Protein = function(DF = DF_p, Covariates_u = F ){
  Association_proteomics = tibble()
  for (Protein in  colnames(DF) ){
    if (Protein %in% c("ID", "Sex", "Age", "CMV_prediction")){ next }
    if (Covariates_u == T){
      Formula = as.formula( paste0("`", Protein, "` ~ Age + Sex + CMV_prediction" ) )
    } else { 
      Formula = as.formula( paste0("`", Protein, "` ~ CMV_prediction" ) )
    }
    lm(Formula, DF) %>% summary() -> temp_result
    temp_result$coefficients %>% as_tibble() %>% tail(1) %>%   mutate(Protein = Protein) -> temp_result
    rbind(Association_proteomics, temp_result) -> Association_proteomics
  }
  return(Association_proteomics)
}
Get_Permutation_FDR = function(Assocation_proteomics,DF, Permutations=100, Use_covariate=F){
  set.seed(87)
  N_perm = Permutations
  Permuted_ps = c()
  for (Permutation in 1:N_perm){
    Permuted_cmv = sample(DF$CMV_prediction, replace=F, size = dim(DF)[1] ) 
     Permuted = Assoc_Protein(mutate(DF, CMV_prediction=Permuted_cmv ), Covariates_u=Use_covariate)
    Permuted_ps = c(Permuted_ps, Permuted$`Pr(>|t|)`)
  }

  FDR = Compute_FDR2(Association_proteomics$`Pr(>|t|)`, Permuted_ps , Permutations)
  Association_proteomics %>% mutate(FDR_permutation = FDR) %>% arrange(`Pr(>|t|)`) -> Association_proteomics
  return(Association_proteomics)
}

#Assocations
Association_proteomics = Assoc_Protein(DF_p_uncorrected, Covariates_u = T   )
#Get FDR
Association_proteomics = Get_Permutation_FDR(Association_proteomics,DF_p_uncorrected, Permutations = 1000)
#Plotting volcano plot
Association_proteomics %>% ggplot(aes(x=Estimate, y=-log10(`Pr(>|t|)` ), col = `Pr(>|t|)`<0.05 )) + geom_point() +theme_bw() + scale_color_discrete(type = met.brewer(name="Klimt", n=2,type="discrete") )
print(Association_proteomics)

write_tsv(Association_proteomics, paste0(Location_dir, "Results/CMV_Olink.tsv"))

```
No significant findings with Olink proteins

Are there Age-dependent differences in CMV effects on proteomics?

Age different effects of CMV in protein levels
```{r Interaction age and CMV }
#Add variable to dataset with different age bins
Covariates %>% mutate(Age_group = ifelse(Age < 30, "Twenties", ifelse(Age<40, "Thirties", ifelse(Age<50, "Fourties", ifelse(Age<60, "Fifties", ifelse(Age<70, "Sixties", ifelse(Age<80, "Seventies", NA) ) ) )  )  ) ) -> Covariates
  Covariates %>% mutate(Age_group = factor(Age_group, levels = c("Twenties", "Thirties", "Fourties", "Fifties", "Sixties", "Seventies" ) ) ) -> Covariates
#Add second variable indicating if young (twenties) or elder (seventies)
Covariates %>% mutate(Age_group2 = ifelse(Age_group %in% c("Twenties", "Seventies"), as.factor(Age_group), NA ) ) -> Covariates



#Function for interaction assocaiton
Interaction_age = function(DF_test){
  Age_CMV = tibble()
  N = colnames(DF_test)[colnames(DF_test) %in% colnames(Proteomics) ]
  for (i in N ){
    if (i == "ID"){ next }
    Model = paste0("`",i, "` ~ Sex + Age_group2*CMV_prediction")
    Model0 = paste0("`", i, "` ~ Sex + Age_group2 + CMV_prediction")
    lm(Model, DF_test) -> Model1
    lm(Model0, DF_test) -> Model0
    anova(Model0, Model1) -> A
    tibble(Protein=i , P=  A$`Pr(>F)`) %>% rbind(Age_CMV, . ) ->  Age_CMV
  }
  return(Age_CMV)
}

set.seed(78789)
Age_CMV = Interaction_age(left_join(Proteomics, Covariates )) %>% drop_na() 
#Perfomr permutations
Permutation_results = tibble()
N_perm = 1000
for (i in seq(N_perm)){
  Covariates_perm = Covariates 
  
  Covariates_perm$CMV_prediction =  sample(Covariates_perm$CMV_prediction, dim(Covariates_perm)[1], replace=T )
  Covariates_perm$Age_group2 =  sample(Covariates_perm$Age_group2, dim(Covariates_perm)[1], replace=T )
  
  Interaction_age(left_join(Proteomics, Covariates_perm, by="ID" ))  %>% drop_na()  %>% rbind(Permutation_results, . ) -> Permutation_results
}
FDRs = Compute_FDR2(Age_CMV$P, Permutation_results , N_perm)
Age_CMV %>% mutate(FDR_p = FDRs) %>% arrange(P) -> Age_CMV

#Get effect size for the signifcant assocaiton
lm(`PSP-D` ~ Sex+as.factor(Age_group2)*CMV_prediction, left_join(Proteomics, Covariates ) ) %>% summary()

#Plot
 left_join(Proteomics, Covariates ) %>% filter(!is.na(CMV_prediction)) %>% filter(! is.na(Age_group)) %>% ggplot( aes(x=as.factor(CMV_prediction) ,y= as.numeric(`PSP-D`)  ) )  + geom_boxplot(outlier.shape = NA) + ggforce::geom_sina(alpha=0.3)  + facet_wrap(~Age_group) + theme_bw()
 left_join(Proteomics, Covariates ) %>% lm(`PSP-D` ~ Sex + Age_group*CMV_prediction, .  ) %>% summary()

write_tsv(Age_CMV, paste0(Location_dir, "Results/CMV_age_interaction_Olink.tsv"))

```


Next, we will investigate the reasons Females have shorter telomeres. This might be related to a stronger response to CMV. Thus we will check: 1. Do females have more proinflammatory proteins than males when infected?  2. Do females have higher antibody titters?


Females TL are more affected by CMV. We wonder if at the cytokine level we also see different effects in females.

```{r Interaction effect of CMV and cytokines levels}


Interaction_association = function(For_association_cytokines){
  Interaction_results = tibble()
  for (Cyto in colnames(dplyr::select(Cytokines, -ID ))){
    Model = paste0("`",Cyto ,"` ~ CMV_prediction + Age + Sex + CMV_prediction:Sex")
    lm(Model, For_association_cytokines) %>% summary() -> Interaction_results_temp
    Interaction_results_temp$coefficients %>% as.data.frame() %>% rownames_to_column("Effect") %>% mutate(Cytokine = Cyto) %>% rbind(Interaction_results, . ) -> Interaction_results 
  }
  return(Interaction_results)
}

#1. 
left_join(Covariates, Cytokines) -> For_association_cytokines
Interaction_association(For_association_cytokines) ->  Interaction_results
#2. Ranking cytokine levels
For_association_cytokines_rank  = left_join(Covariates, Cytokines %>% dplyr::select(-ID) %>% apply(2, InvRank) %>% as_tibble() %>% mutate(ID = Cytokines$ID) )
Interaction_association( For_association_cytokines_rank ) -> Interaction_results_rank
#3. Permutation-based FDR
Do_permutations = function(Interaction_results, shuffle="CMV_and_sex", N_perm=1000){
  set.seed(8744)
  Result_permutation = tibble()
  for ( i in seq(1:N_perm) ){
    print(paste0("Permutation: ", i ))
    Covariates_perm= Covariates
    if (shuffle == "CMV") {
      sample(c(1,0), replace=T , 1448) -> NID
      Covariates_perm$CMV_prediction = NID
    } else if (shuffle == "CMV_and_sex") {
      sample(Covariates_perm$CMV_prediction, replace=T , 1448) -> NID
      sample(Covariates_perm$Sex, replace=T , 1448) -> NID2
      Covariates_perm$CMV_prediction = NID
      Covariates_perm$Sex = NID2
    }
    Interaction_association(left_join(Covariates_perm, Cytokines, by="ID")) %>% rbind(Result_permutation, . ) -> Result_permutation
    
  }
  Result_permutation %>% filter(grepl(":", Effect)) -> Result_permutation
  #Result_permutation %>% ggplot(aes(x=`Pr(>|t|)` )) + geom_histogram() + theme_bw() + facet_wrap(~Cytokine) 
  
  # calculate FDR for each observed P-value
  FDRs = Compute_FDR2(Interaction_results$`Pr(>|t|)`, Result_permutation$`Pr(>|t|)` , N_perm)
  return(Interaction_results %>% mutate(FDR_p = FDRs))
}  
  
#Results using the rank-based approach  
Interaction_results_rank %>% filter(grepl(":", Effect)) -> Interaction_results_rank
Do_permutations(Interaction_results_rank, N_perm = 1000 )  -> Interaction_results_rank
Interaction_results_rank %>% mutate(Model = "Rank") %>% as_tibble() %>% arrange(`Pr(>|t|)` ) -> Interaction_results_rank
#Results without rank          
Interaction_results %>% filter(grepl(":", Effect)) -> Interaction_results
Do_permutations(Interaction_results ) -> Interaction_results
Interaction_results %>% mutate(Model = "Rank") %>% as_tibble() %>% arrange(`Pr(>|t|)` )  -> Interaction_results
#Combine both
left_join(Interaction_results_rank %>% select(-Model),Interaction_results %>% select(-Model) , by=c("Effect","Cytokine"), suffix=c("_rankedCytokine","")  ) %>% arrange(`Pr(>|t|)_rankedCytokine`) -> For_Save
rbind(Interaction_results_rank,Interaction_results )  -> Interaction_results_all
#Save
For_Save  %>% write_tsv(. , path = paste0(Location_dir, "Results/CMV_cytokine_sex_interaction.tsv"))

#Volcano plot
Interaction_results_all %>% filter(Model=="Rank") %>% ggplot(aes(x=Estimate, y=-log10(`Pr(>|t|)`), col = FDR_p<0.05 )) + geom_point() + theme_bw() + geom_vline(xintercept = 0) + geom_hline(yintercept = -log10(0.05), linetype=2) +scale_color_manual(values = c(met.brewer(name="Klimt", n=2,type="discrete")) ) #+ facet_wrap(~Model, scales = "free")

#Plots sex dependent assocaiton
#Prepare data
For_association_cytokines %>% mutate(Sex_and_status = ifelse(CMV_prediction==1 & Sex ==2, "Female positive", ifelse(CMV_prediction==1 & Sex==1, "Male positive", ifelse(CMV_prediction==0 & Sex==2, "Female negative", "Male negative") ) ) ) -> For_plot
left_join(Covariates, Cytokines %>% select(-ID) %>% apply(2, InvRank) %>% as_tibble() %>% mutate(ID = Cytokines$ID) )  %>% mutate(Sex_and_status = ifelse(CMV_prediction==1 & Sex ==2, "Female positive", ifelse(CMV_prediction==1 & Sex==1, "Male positive", ifelse(CMV_prediction==0 & Sex==2, "Female negative", "Male negative") ) ) ) -> For_plot2
#plot
For_plot %>% filter(!is.na(Sex_and_status)) %>% filter(`IL.1.Beta` >0.5) %>%  ggplot(aes(x= as.factor(Sex_and_status) , y = log10(`IL.1.Beta`) )) + geom_boxplot() + theme_bw() + ggforce::geom_sina()
For_plot2 %>% filter(!is.na(Sex_and_status)) %>% filter(`IL.6` > 0) %>%  ggplot(aes(x= as.factor(Sex_and_status) , y = `IL.6`) ) + geom_boxplot() + theme_bw() + ggforce::geom_sina()+ coord_flip() +  theme(axis.title.y=element_blank() ) -> Il6_int
For_plot2 %>% filter(!is.na(Sex_and_status)) %>% filter(`IL.12P70` > 0) %>%  ggplot(aes(x= as.factor(Sex_and_status) , y = `IL.12P70`) ) + geom_boxplot() + theme_bw() + ggforce::geom_sina() + coord_flip() +  theme(axis.title.y=element_blank() )  -> IL12P70_int

Il6_int | IL12P70_int -> cytokine_plots
ggsave( paste0(Location_dir,"Results/Figures/Sex_cytokine_CMV.pdf"), cytokine_plots)


For_plot2 %>% filter(`IL.12P70` > 0) %>% lm(`IL.12P70` ~ Age + CMV_prediction*Sex, . ) %>% summary()


lm(`IL.12P70` ~ Age + CMV_prediction ,For_association_cytokines_rank %>% filter(Sex==1)  ) %>% summary()
lm(`IL.12P70` ~ Age + CMV_prediction ,For_association_cytokines_rank %>% filter(Sex==2)  ) %>% summary()

lm(`IL.6` ~ Age + CMV_prediction ,For_association_cytokines_rank %>% filter(Sex==1) ) %>% summary()
lm(`IL.6` ~ Age + CMV_prediction ,For_association_cytokines_rank %>% filter(Sex==2) ) %>% summary()

```

There are two significant results if rank, none if not ranked.


Do females also show higher levels of CMV titters than males?

```{r Differences in titters between sexes}
  All_alr = read_rds( paste0(Location_dir,"Data/Quantifiaction_PhIPseq_alr.rds") )
  cmv_peptides <- read_tsv(paste0(Location_dir, "Results/CMV_peptides.tsv"))
  cmv_peptides <- cmv_peptides$Peptide


  All_alr %>% filter(grepl("LLD", ID)) %>% filter(! grepl("_F", ID) ) -> All_alr
  All_alr %>% filter(ID %in% filter(Covariates, CMV_prediction == 1)$ID ) %>% dplyr::select(c("ID", cmv_peptides)) -> Titters
  
  Titters %>% left_join(Covariates %>% select(ID, Sex, Age )) -> Titters_test
  Association_titters = tibble()
  for (Peptide in colnames(Titters)){
    if (Peptide == "ID"){ next }
    Model = paste0(Peptide, " ~ Sex + Age")
    lm(Model, Titters_test) %>% summary() -> Model
    Model$coefficients %>% as.data.frame() %>% rownames_to_column("Value") %>% filter(Value == "Sex") %>% dplyr::select(-Value) %>% as_tibble() %>% mutate(Peptide = Peptide) %>% rbind(Association_titters, . ) -> Association_titters
  }
  write_tsv(Association_titters, paste0(Location_dir,"Results/CMV_titters_vs_female.tsv"))
  Association_titters %>% arrange(`Pr(>|t|)`) %>% ggplot(aes(x=Estimate, y=-log10(`Pr(>|t|)`) )) + theme_bw() + geom_point() + xlab("Female effect on CMV antibody titter\n in CMV positive individuals") + xlim(-0.30, 0.3) + geom_vline(xintercept = 0,linetype="dotdash") + geom_hline(yintercept = -log10(0.05), linetype= "dotdash") -> Figure_titer
  ggsave(paste0(Location_dir,"Results/Figures/SexTiterDifferences_CMVPositive.pdf"), Figure_titer)
  
  Association_titters %>% arrange(`Pr(>|t|)`) %>% write_tsv(. , path = paste0(Location_dir, "Results/CMV_vs_Titters.tsv"))
  Figure_titer
```

Replication of sex differences in titters

```{r Replication CMV-titters in females: 300BCG}
#CMV measurements. 
read_excel(  paste0(Location_dir,"Data/Replication/300BCG/2019 300BCG serologies_for_sergio.xlsx")) -> Info_serology_300bcg #Laurent Gillet’s lab measurements
read_excel( paste0(Location_dir,"Data/Replication/300BCG/CMV IgG index 300BCG.xlsx"), sheet= "CMV_for_R") -> Info_serology_300bcg2 #JCI Olink paper (‘PMID: 32692728’)
load( paste0(Location_dir,"Data/Replication/300BCG/data.Robj"), verbose=T) #Metadata from the cohort. Downloaded from 'PMID: 32692728’ GitLab's
obj[[2]] %>% as_tibble() -> metadata_300bcg


metadata_300bcg %>% rename(PatientID = `Patient,ID`) %>% left_join(Info_serology_300bcg) -> Replication_300bcg
left_join(Replication_300bcg, Info_serology_300bcg2 ) -> Replication_300bcg2

#Check age and sex
Replication_300bcg %>% ggplot(aes(x=Age, fill=CMV )) + geom_histogram() + theme_bw()
Replication_300bcg$Gender %>% table()



#Check that IDs are good by checking age and sex assocasiations
Replication_300bcg %>% filter(CMV %in% c("+", "-") ) %>%  glm( as.factor(CMV) ~ Gender + Age, . , family=binomial()) %>% summary() # relative assocaiton, not too strong


#Check for titter assocaitons with females
Replication_300bcg %>% filter(CMV == "+" ) %>% lm(`CMV ELISA` ~ Gender, . ) %>% summary() #Negative but not significant
#This assocation is not significant at all. However, data from 32692728 is more promising
Replication_300bcg2 %>% filter(CMV_status == "Positive" ) %>%  mutate(CMV_IgG_IU_ml = as.numeric(CMV_IgG_IU_ml))  %>% lm( CMV_IgG_IU_ml ~ Gender + Age , . )  %>% summary()  #IgG titters
Replication_300bcg2 %>% filter(CMV_status == "Positive" ) %>%  mutate(CMV_IgG_Index = as.numeric(CMV_IgG_Index))  %>% lm( CMV_IgG_Index ~ Gender + Age , . )  %>% summary() #IgG index


#PLOTS 
Replication_300bcg2 %>% filter(CMV_status == "Positive" ) %>% ggplot(aes(x=Gender, y = as.numeric(CMV_IgG_IU_ml))) + geom_boxplot() + theme_bw() + geom_sina() + coord_flip() #Seems like there is a an upper limit in Units per ml
Replication_300bcg2 %>% filter(CMV_status == "Positive" ) %>% ggplot(aes(x=Gender, y = as.numeric(`CMV ELISA`))) + geom_boxplot() + theme_bw() + geom_sina() + coord_flip()
Replication_300bcg2 %>% filter(CMV_status == "Positive" ) %>% ggplot(aes(x=Gender, y = as.numeric(`CMV_IgG_Index`))) + geom_boxplot() + theme_bw() + geom_sina() + coord_flip()




```


Replciation in 500FG?

